Comunicações Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC

### 2015-05-06 tese/section04.tex
english syntax for section 1, webrtc stack figure
quase a roçar as 16 paginas
added references, almost concluding syntax checking for section 3

 	Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity. Technologies like this are appearing along this days, but they couln't be possible years ago because technology was limited or unavailable. Although todays technology is limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.

@@

	If, for instance, one whants to rewind a real time video, recordings will be needed from who is streaming the video.

  \ac{RTP}\footnote{rfc3550} is used for streaming audio and video over \ac{IP}, the multimedia content is transported on the payload of \ac{RTP} messages, \ac{RTP} contains headers for payload indentification. \ac{RTP} is independent from its payload type, allowing to transport any kind of encoded multimedia. A sequence number is used for sorting received packets.

@@

  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \footnote{rfc3551} which lists the payload encodings and compression algorithms. This profile also assigns a name to each encoding which may be used other protocols like \ac{SDP}.

@@

  Both \ac{SRTP} and \ac{SRTCP} uses \ac{AES} by default, which is symmetric-key algorithm for data encryption. Each packet is encrypted using a distinct keystream, otherwise using a single keystream with \ac{AES} on \ac{CBC} would make impossible to recover from packet loss.

  wo keystream generators for \ac{AES} were defined: \textit{Segmented Integer Counter Mode} and \textit{f8-mode}. If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those keystream generators and it's fixed for each packet.

@@

\begin{figure}[H]
	\begin{center}
		\centering
		\includegraphics[width=0.35\textwidth]{figures/media_types.png}

	\caption{Media Types}
	\end{center}
\end{figure}

	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuos, the second one describes it as interactive or non-interactive. A discrete media is caracterized by its type not depending from time, as continuous media depending from it. Interactive media is caracterized by its state being changed by external events such as user interactions.

	For example, an image is non-interctive and discrete, for instance, a video is continuos and non-interactive. A simple colaborative editor with just text is interactive and discrete. An animation that changes in function of user behaviour is interactive and continuous.

	Streaming protocols like \ac {RTP} were designed for continuous and non-interactive media types, such as audio and video. Discrete and non-interactive media don't need to streamed through \ac{RTP} because they don't change with. For example if an image appears in a specific time interval, just the \ac{HTML} or javascript that will reference the image must be streamed, the image itself is then transfered through \ac{HTTP}.

@@

	Everytime an event is proccessed on one of the endpoints, both sender and receivers state must stay synchronized, otherwise events may behave differently.

	To achieve synchronization of interactive data most packets have three types: \textit{State}, \textit{Delta-State} and \textit{Event}. State packet defines environment's complete state. Delta-State packets transports just the piece of state that changed. Event packets informs that an event ocurred over the interactive media.

	An \ac{RTP} recorder can have two operation modes, recording or playback. Traditional \ac{RTP} players can do random access, in contrast, interactive \ac{RTP} players must restore the environment and context at a given time. The environment is the initial state, so we can call it a non-interactive descrete media and handle it over \ac{HTTP}. After the receiver has received the environment, it should calculate the state at the given time.

@@

	Multiparty video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants. Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.

	Jitsi Video Bridge \footnote{jitsi VB} receives one stream from every participant on a conference, either from a jitsi client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends. Although all the participants need to download all the streams from Jitsi Video Bridge Server, typically download rates are much bigger than upload rates, making this solution more feasible.

	Jitsi Video Bridge uses \ac{XMPP} as a signaling protocol and its colibri extension (XEP-0340) to reserve channels for video transmission. Although this choice for signalling protocol, Jitsi Video Bridge also supports \ac{SIP} and \ac{Jingle}.

>>>>

\subsection{Extending collaboration tools with time manipulation}

 	Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity. Technologies like this are appearing along this days, but they couldn't be possible years ago because technology was limited or unavailable. Although todays technology is limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.

@@

	If, for instance, one wants to rewind a real time video, recordings will be needed from who is streaming the video.

  \ac{RTP}\cite{rfc3550} is used for streaming audio and video over \ac{IP}, the multimedia content is transported on the payload of \ac{RTP} messages, \ac{RTP} contains headers for payload identification. \ac{RTP} is independent from its payload type, allowing to transport any kind of encoded multimedia. A sequence number is used for sorting received packets.

@@

  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \cite{rfc3551} which lists the payload encodings and compression algorithms. This profile also assigns a name to each encoding which may be used other protocols like \ac{SDP}.

@@

  Both \ac{SRTP} and \ac{SRTCP} uses \ac{AES} by default, which is symmetric-key algorithm for data encryption. Each packet is encrypted using a distinct key-stream, otherwise using a single key-stream with \ac{AES} on \ac{CBC} would make impossible to recover from packet loss.

  Two key-stream generators for \ac{AES} were defined: \textit{Segmented Integer Counter Mode} and \textit{f8-mode}. If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those key-stream generators and it's fixed for each packet.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{figures/media_types.png}
	\caption{Media Types}
\end{figure}

	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive. A discrete media is characterized by its type not depending from time, as continuous media depending from it. Interactive media is characterized by its state being changed by external events such as user interactions.

	For example, an image is non-interactive and discrete, for instance, a video is continuous and non-interactive. A simple collaborative editor with just text is interactive and discrete. An animation that changes in function of user behavior is interactive and continuous.

	Streaming protocols like \ac {RTP} were designed for continuous and non-interactive media types, such as audio and video. Discrete and non-interactive media don't need to streamed through \ac{RTP} because they don't change with. For example if an image appears in a specific time interval, just the \ac{HTML} or \textit{JavaScript} that will reference the image must be streamed, the image itself is then transfered through \ac{HTTP}.

@@

	Every time an event is processed on one of the endpoints, both sender and receivers state must stay synchronized, otherwise events may behave differently.

	To achieve synchronization of interactive data most packets have three types: \textit{State}, \textit{Delta-State} and \textit{Event}. State packet defines environment's complete state. Delta-State packets transports just the piece of state that changed. Event packets informs that an event occurred over the interactive media.

	An \ac{RTP} recorder can have two operation modes, recording or playback. Traditional \ac{RTP} players can do random access, in contrast, interactive \ac{RTP} players must restore the environment and context at a given time. The environment is the initial state, so we can call it a non-interactive discrete media and handle it over \ac{HTTP}. After the receiver has received the environment, it should calculate the state at the given time.

@@

	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants. Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.

	Jitsi Video Bridge \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a jitsi client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends. Although all the participants need to download all the streams from Jitsi Video Bridge Server, typically download rates are much bigger than upload rates, making this solution more feasible.

	Jitsi Video Bridge uses \ac{XMPP} as a signaling protocol and its colibri extension \cite{xep0340} to reserve channels for video transmission. Although this choice for signaling protocol, Jitsi Video Bridge also supports \ac{SIP}.
